---
title: "Group_5_Analysis"
output: html_document
---

#Load Libraries
```{r}
#install.packages(sf)

library(tidyverse)
library(lubridate)
library(sf)
options(scipen = 999)

```

#Load Data
```{r}
# Street Racing Fatality Data from 2000 to 2017

streetracing <- read_csv("Data/street-racing-fatalities.csv")

# More fatality data

fatalities_by_region <- read_csv("Data/fatalities.csv")

# More

citypop <- st_read("")

```

#Sentence 1: [Valentina D’Alessandro] was one of at least 179 people who have died in Los Angeles County since 2000 in accidents where street racing was suspected, according to a Los Angeles Times analysis of coroner’s records, police reports and media accounts from 2000 to 2017.


#Sentence 2: The deadliest year on record was 2007, with 18 fatal crashes. After a period with relatively few recorded deaths, the count grew in recent years, with 15 fatalities in 2015, 11 in 2016 and 12 deaths in 2017, the Times analysis shows.

```{r}
deaths_by_year <- streetracing%>%
  mutate(year = year(DeathDateTime))%>%
  select(year, everything())
 
deaths_by_year

```

```{r}
none <- streetracing %>%
  anti_join(fatalities_by_region,by = c("CaseNum"="case_number") )

none

```

#Sentence 3: The dead were overwhelmingly male and young: More than half were 21 or younger, including two children, ages 4 and 8, killed along with their mother. Slightly less than half of the victims— 47% — were behind the wheel. The rest were either passengers in the speeding vehicles, spectators or people simply walking on a sidewalk or driving home.


# Sentence 4: Deaths have occurred all across L.A. County, with the biggest concentration in the small city of Commerce. Eleven have died on the small, industrial city’s streets since 2000.

```{r}
#Selecting columns and renaming to fatalities_truncated to make data concise.

fatalities_truncated<-fatalities_by_region%>%
  select(name, city, death_datetime, death_year)

#Glimpse findings below.
glimpse(fatalities_truncated)
```

```{r}
#Filtering to find deaths in Commerce City.

commerce_city_deaths <- fatalities_truncated%>%
  filter(name == "Commerce")

#show findings below
commerce_city_deaths

#Q: How many deaths occurred in Commerce City since 2000?
#A: 11
```

```{r}
death_by_city <- fatalities_truncated%>%
  group_by(name)%>%
  summarise(city_deaths = n())%>%
  arrange(desc(city_deaths))

death_by_city

```

```{r}
death_by_city_pop <- death_by_city %>%
  inner_join(citypop) %>%
  mutate(deaths_per_capita = (city_deaths/population*10000)) %>%
  filter(population > 1000) %>%
  arrange(desc(deaths_per_capita))

death_by_city_pop

